---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "man/figures/README-", out.width = "100%")
```

# `rTVGL`: Time-Varying Graphical LASSO for R

This R package implements the time-varying graphical LASSO method introduced by Hallac et al. (2017). The ADMM algorithm is adapted from the Python package `regain` (Tomasi et al., 2018).

## Installation

The `rTVGL` package can be installed using the following code:

```{r, eval = FALSE}
if(!require("devtools", quietly = TRUE)) {
  install.packages("devtools")
}
devtools::install_github("THautamaki/rTVGL")
```

## Example of usage

This is a minimum working example. Load the package and define variables used in the data simulation.

```{r}
library(rTVGL)
n <- 50         # Number of observations per time point (sample size)
p <- 100        # Number of variables
n_tp <- 20      # Number of time points
step_size <- 5  # Number of time points between changes in the network structure
cps <- seq(step_size, n_tp - step_size, step_size)  # Vector of actual change points
number_add_del <- 10  # Number of edges been deleted and added in the change point
```

Generate simulated data using function `generate_timeseries_network_data`.

```{r, message = FALSE, results = 'hide'}
sim <- generate_timeseries_network_data(n = n, p = p, n_timepoints = n_tp, change_points = cps,
                                        n_add_del = number_add_del, seed_net = 20260511,
                                        seed_data = 20260511)
```

Run the TVGLASSO method using the `tvgl` function and $l_1$ penalty over the time.

```{r}
results <- tvgl(sim$datasets[[1]], lambda = 0.25, beta = 0.25, penalty_type = "l1")
```

Calculate and print the confusion matrix for the first timepoint.

```{r}
(cm <- conf_matrix(sim$Thetas[,,1], results$Theta_ests[,,1]))
```

Calculate and print some performance scores.

```{r}
round(calculate_scores(cm)[, c("MCC", "F1", "TPR", "FDR")], 4)
```

Plot the true simulated network and estimated network using the R package `igraph`.

```{r, message = FALSE}
if(!require("igraph", quietly = TRUE)) {
  install.packages("igraph")
}
library(igraph)
```

Generate graph objects using `graph_from_adjacency_matrix` function from `igraph` package.

```{r}
true_network <- graph_from_adjacency_matrix(sim$Thetas[,,1], mode = "undirected", diag = F)
est_network <- graph_from_adjacency_matrix(results$Theta_ests[,,1], mode = "undirected", diag = F)
```

Create coordinates for the nodes of the network. Use the same node placement for both plots.

```{r}
set.seed(10)
coords <- layout_with_fr(true_network)
```

Finally, plot both the true and estimated networks.

```{r network_plots, fig.dpi = 300}
par(mfrow = c(1,2), mar = c(0.1, 0.1, 1, 0.1))
plot(true_network, layout = coords, edge.width = 2, vertex.size = 4, vertex.label = NA,
     main = "True network")
plot(est_network, layout = coords, edge.width = 2, vertex.size = 4, vertex.label = NA,
     main = "Estimated network")
```

## References

-   Hallac, D., Park, Y., Boyd, S., & Leskovec, J. (2017). Network inference via the time-varying graphical lasso. In *Proceedings of the 23rd ACM SIGKDD international conference on knowledge discovery and data mining* (pp. 205-213).
-   Tomasi, F., Tozzo, V., Salzo, S., & Verri, A. (2018). Latent variable time-varying network inference. In *Proceedings of the 24th ACM SIGKDD International Conference on Knowledge Discovery & Data Mining* (pp. 2338-2346).
